Clustering Beat-Chroma Patterns\\ in a Large Music Database

### 2010-03-24 Paper/Paper.tex
submittable... almost, missing new frequency figure

One major issue in working with small patches concerns temporal
alignment. Learning the same motifs with different offset
(of a few seconds, or audio frames) is redundant. We take advantage
of rythmic knowledge about the song to align patches with beats.
We can further group beats into bars, meaningful music segments.
The use of the Echo Nest API \cite{EchoNest} for that purpose is 
explained in Subsection \ref{ssec:beatpatch}. Patches built around
those temporal landmarks should capture music motifs more easily.

Prior work using small patches of chroma features includes
the ``shingles'' of \cite{Casey2007}, which they used to
identify ``remixes'' i.e. music based on some of the same
underlying instrument tracks, as well as for matching
performances of Mazurkas \cite{Casey2008}.  That work,
however, was not concerned with extracting the deeper
common patterns underlying different pieces (and did not
use beat- or bar-synchronous features).  Earlier work in
beat-synchronous analysis includes \cite{Bartsch2001},
who looked for repeated patterns within single songs to
identify the chorus, and \cite{Ellis2007a}, who
cross-correlated beat-chroma matrices to match cover
versions of pop music tracks.  None of these works
examined or interpreted the content of the chroma
matrices in any detail.  In contrast, here we hope to
develop a codebook whose entries are of interest
in their own right.

@@

\subsection{Basic Properties}
This section present some fundamental results of the clustering algorithm.
None of the result appear surprising,  but they are important for 
reproducing the experiments.
\begin{itemize}
\item Encoding performance improves with increasing training data (Figure
\ref{fig:data_sizes}). Regarding the covergence of the performance,
we can imagine using $500$ or $1$K samples per codeword to learn if the
training set size is problematic.
\item Encoding performance improves with increasing codebook size
(Table \ref{tab:cbsize}). The increase seems significant, and
increasing the size of the codebook in an efficient way
is definitely part of our future work.
\item Larger patterns are more difficult to encode, thus requiring
larger codebooks. See Figure \ref{fig:perbeat}. The increase is steeper
until $4$ beats which usually corresponds to one bar.
\end{itemize}

@@

Another interesting visualization uses LLE\footnote{implementation:
  \url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}
algorithm \cite{Roweis2000} to display codewords arranged in
coherent groups. See Figure \ref{fig:lle}. FOr instance, we see
``wideband'' patterns grouping to the right.

@@

We apply our codebook to the artist recognition task.
Our data set encompass $1402$ songs from $20$ artists, 
mostly rock and pop of different subgenres. Results exist in the
litterature for that set \cite{Ellis2007}. The best accuracy is
$59\%$, whereas using only chroma as a representation yields an
accuracy of $33\%$. The latter results is obtained through a
Gaussian mixture model.

@@

Artists like Metallica are characterized by ``wideband'' patterns -
featured muddled due to noisy acoustics - lots of transients,
distortion, etc.  (not by cleaned up by EN's feature analysis)

>>>>

One major issue in working with small patches concerns temporal
alignment. Learning the same motifs with different offset
(of a few seconds, or audio frames) is redundant. We take advantage
of rhythmic knowledge about the song to align patches with beats.
We can further group beats into bars, meaningful music segments.
The use of the Echo Nest API \cite{EchoNest} for that purpose is 
explained in Subsection \ref{ssec:beatpatch}. Patches built around
those temporal landmarks should capture music motifs more easily.

Prior work using small patches of chroma features includes
the ``shingles'' of \cite{Casey2007}, which they used to
identify ``remixes'' i.e. music based on some of the same
underlying instrument tracks, as well as for matching
performances of Mazurkas \cite{Casey2008}.  That work,
however, was not concerned with extracting the deeper
common patterns underlying different pieces (and did not
use either beat- or bar-synchronous features).  Earlier work in
beat-synchronous analysis includes \cite{Bartsch2001},
who looked for repeated patterns within single songs to
identify the chorus, and \cite{Ellis2007a}, who
cross-correlated beat-chroma matrices to match cover
versions of pop music tracks.  None of these works
examined or interpreted the content of the chroma
matrices in any detail.  In contrast, here we hope to
develop a codebook whose entries are of interest
in their own right.

@@

\subsection{Basic Properties}
This section present some fundamental results of the clustering algorithm.
None of the result appear surprising,  but they are important for 
reproducing the experiments.
\begin{itemize}
\item Encoding performance improves with increasing training data (Figure
\ref{fig:data_sizes}). Regarding the convergence of the performance,
we can imagine using $500$ or $1$K samples per codeword to learn if the
training set size is problematic.
\item Encoding performance improves with increasing codebook size
(Table \ref{tab:cbsize}). The increase seems significant, and
increasing the size of the codebook in an efficient way
is definitely part of our future work.
\item Larger patterns are more difficult to encode, thus requiring
larger codebooks. See Figure \ref{fig:perbeat}. The increase is steeper
until $4$ beats which usually corresponds to one bar.
\end{itemize}

@@

Another interesting visualization uses LLE\footnote{implementation:
  \url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}
algorithm \cite{Roweis2000} to display codewords arranged in
coherent groups. See Figure \ref{fig:lle}. For instance, we see
``wideband'' patterns grouping to the right.

@@

We apply our codebook to the artist recognition task.
Our data set encompass $1402$ songs from $20$ artists, 
mostly rock and pop of different subgenres. Results exist in the
literature for that set \cite{Ellis2007}. The best accuracy is
$59\%$, whereas using only chroma as a representation yields an
accuracy of $33\%$. The latter results is obtained through a
Gaussian mixture model.

@@

Artists like Metallica are characterized by ``wideband'' patterns -
featured muddled due to noisy acoustics - lots of transients,
distortion, etc, which can be seen in Figure \ref{fig:metallica2}.
%FIXME: ron, what di you mean: (not by cleaned up by EN's feature analysis)
